Clique and Constraint Models for Maximum Common (Connected) Subgraph Problems

### 2016-04-20 paper/paper.tex
Wording

In \cref{algorithm:mccis} we present a clique-inspired algorithm which finds a maximum common
connected induced subgraph isomorphism via an association graph. If the additional branching
restrictions are removed, the core of the algorithm is the same as the ``MCSa1'' clique algorithm
used in the previous section (and we refer the reader to the previously cited works for
implementation details on how to use bitsets and other data structures to implement the colouring
stage with very low constant factors). The approach we describe here differs considerably from that
of Koch \cite{DBLP:journals/tcs/Koch01} and Vismara and Valery \cite{DBLP:conf/mco/VismaraV08}:
these earlier approaches worked by classifying labels in the association graph based upon whether a
common vertex is shared, and then constructing cliques with particular edge properties---this is
harder to integrate with a strong bound function.

We begin by building the association graph (\lineref{buildassoc}). The main part of the algorithm
then works by building up candidate cliques in the $\mathit{solution}$ variable, by recursive calls
to the $\mathit{search}$ procedure---starting from the empty set (\lineref{firstsearch}), each
recursive subcall adds one vertex to $\mathit{solution}$ (\lineref{addv}) in such a way that
$\mathit{solution}$ is always a clique which corresponds to a connected common subgraph. The
$\mathit{remaining}$ set contains the set of vertices which are adjacent to every vertex in
$\mathit{solution}$, and which have not yet been accepted or rejected (and so initially it contains
every vertex). The main loops in the $\FuncSty{search}$ procedure
(\twolinesref{outerloop}{innerloop}) have the effect of iterating over each vertex in this set in a
particular order---each vertex $v$ is selected in turn, and then a recursive call is made to
consider the effects of including $v$ in $\mathit{solution}$ (\lineref{recursesearch}), followed by
the next iteration where $v$ is instead rejected. When $v$ is accepted, we add it to the new
$\mathit{solution'}$ (\lineref{addvtosolution}), and create a new $\mathit{remaining'}$ containing
only the vertices in $\mathit{remaining}$ which are adjacent to $v$ (\lineref{filterremaining}).

The $\mathit{branchable}$ set contains the set of association graph vertices which correspond to
vertices adjacent to an already-accepted vertex in the first input graph---in constraint programming terms, it is the
set of assignments which could be made next which maintain connectedness. (Using only one of the two
input graphs is sufficient for correctness, and has the advantage that the $\FuncSty{connected}$
function may be implemented as a simple lookup into a precomputed array which maps each vertex in
the first input graph to a bitset.) At the top of search, this set is empty, and is not used (our
first vertex selection is special, and does not care about connectedness). At subsequent depths, we
may only accept vertices which are in this set, and if no such vertices remain then we return
immediately (\lineref{acceptbranchable}). When recursing, we extend $\mathit{branchable}$ with the
new vertices permitted by our acceptance of the branching $v$ (\lineref{addtobranchable}). Note that
we are assuming that inside the main loops, we encounter every vertex in $\mathit{remaining} \cap
\mathit{connected}$ before any vertex in $\mathit{remaining} \setminus \mathit{connected}$.

>>>>

In \cref{algorithm:mccis} we present a clique-inspired algorithm which finds a maximum common
connected induced subgraph isomorphism via an association graph. If the additional branching
restrictions are removed, the core of the algorithm is the same as the ``MCSa1'' clique algorithm
used in the previous section (and we refer the reader to the previously cited works for
implementation details on how to use bitsets and other data structures to implement the colouring
stage with very low constant factors). The way we extend this algorithm for connectedness differs
considerably from that of Koch \cite{DBLP:journals/tcs/Koch01} and Vismara and Valery
\cite{DBLP:conf/mco/VismaraV08}: these earlier approaches worked by classifying labels in the
association graph based upon whether a common vertex is shared, and then constructing cliques with
particular edge properties---this is harder to integrate with a strong bound function.

Our algorithm begins by building the association graph (\lineref{buildassoc}). The main part of the algorithm
then works by building up candidate cliques in the $\mathit{solution}$ variable, by recursive calls
to the $\mathit{search}$ procedure---starting from the empty set (\lineref{firstsearch}), each
recursive subcall adds one vertex to $\mathit{solution}$ (\lineref{addv}) in such a way that
$\mathit{solution}$ is always a clique which corresponds to a connected common subgraph. The
$\mathit{remaining}$ set contains the set of vertices which are adjacent to every vertex in
$\mathit{solution}$, and which have not yet been accepted or rejected (and so initially it contains
every vertex). The main loops in the $\FuncSty{search}$ procedure
(\twolinesref{outerloop}{innerloop}) have the effect of iterating over each vertex in this set in a
particular order---each vertex $v$ is selected in turn, and then a recursive call is made to
consider the effects of including $v$ in $\mathit{solution}$ (\lineref{recursesearch}), followed by
the next iteration where $v$ is instead rejected. When $v$ is accepted, we add it to the new
$\mathit{solution'}$ (\lineref{addvtosolution}), and create a new $\mathit{remaining'}$ containing
only the vertices in $\mathit{remaining}$ which are adjacent to $v$ (\lineref{filterremaining}).

The $\mathit{branchable}$ set contains the set of association graph vertices which correspond to
vertices adjacent to an already-accepted vertex in the first input graph---in constraint programming terms, it is the
set of assignments which could be made next which maintain connectedness. (Using only one of the two
input graphs is sufficient for correctness, and has the advantage that the connectedness set
may be determined by a simple lookup into a precomputed array which maps each vertex in
the first input graph to a bitset.) At the top of search, this set is empty, and is not used (our
first vertex selection is special, and does not care about connectedness). At subsequent depths, we
may only accept vertices which are in this set, and if no such vertices remain then we return
immediately (\lineref{acceptbranchable}). When recursing, we extend $\mathit{branchable}$ with the
new vertices permitted by our acceptance of the branching $v$ (\lineref{addtobranchable}). Note that
we are assuming that inside the main loops, we encounter every vertex in $\mathit{remaining} \cap
\mathit{connected}$ before any vertex in $\mathit{remaining} \setminus \mathit{connected}$.
